OR Multi Disturbance

### 2025-11-21 OR_model_DisturbPrior.Rmd
in-progress versions of OR model that adds disturbance product priors

---
title: "OR Multi Disturbance"
author: "Michael Dietze"
date: "12/7/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
source("disturbance.R")
source("multi.ens.adj.R")
```

## set up
```{r}
## Load disturbance rate, state&transition, disturbed biomass stats
load("STprior.RData")
colnames(dextract) <- paste0("t",dtime)
dextract[is.na(dextract)] <- 0
d.prior["UN"] = d.prior["UN"] + d.prior[6]                         ## change at runtime (removes <NA>)
d.prior = d.prior[-6]

## Load Calibration
load("fit.RData")
Qsoil = Qs

## Load extracted site timeseries
load("extract.h03v04.RData")
```

@@

## functions
```{r,echo=FALSE,results='hide',message=FALS}
##' Single Site, Single Disturbance
##' @param Y observed data
##' @param R observation error
##' @param Q process error
##' @param Xic state initial condition matrix
##' @param THETA parameter matrix
##' @param dist  disturbance list: p.dist, mu0[3], V0[3,3], alloc soil 
SSSD.model <- function(Y,R,Q,PAR,Xic,THETA,dist){

@@

    ## Disturbance
    d.forecast <- rep(0,ne); d.forecast[1:ceiling(ne*0.1)] <- 1 #rbinom(ne,1,dist$p.dist)             ## prior on disturbance status
    sel <- !as.logical(d.forecast)                ## select undisturbed ensemble members
    if(sum(!sel)>2){
      sel.dist = which(!sel)
      for(j in sel.dist){                      ## run disturbance model
        Nf[j,]  = disturbance.p(Nf[j,],mu0=dist$mu0,V0 = dist$V0,alloc.soil = dist$alloc.soil)
      }
      mufD  = colMeans(Nf[!sel,])
      pfD = solve(var(Nf[!sel,]))#solve(diag(c(0.15,0.5,0.25)^2))
    } 
    if(sum(!sel) <= 2 || class(pfD) == "try-error") {   ## backup values if the ensemble stats fail
      mufD = c(dist$mu0[1], mean(Nf[,2]),  dist$mu0[2])            
      pfD  = solve(diag(c(0.15,0.5,0.25)^2))
    }
    Nf <- tobit(Nf)

    ## priors
    priors <- list(p    = dist$p.dist,         ## disturbance rate
                   mufN = colMeans(Nf[sel,]),    ## Not disturbed forecast means
                   mufD = mufD,   ## Disturbed forecast means
                   pfN  = solve(var(Nf[sel,])),  ## Not disturbed forecast precision
                   pfD  = pfD  ## Disturbed forecast precision
    )
    Forecast[[t]] <- c(list(d.forecast=d.forecast,Nf=Nf),priors)

@@

      ## update parameters & save Analysis posterior
      dat = as.matrix(jdat)
      nsel <- grep(pattern="^n",x=colnames(dat))
      Xa = dat[,nsel]
      ca = dat[,"D"]
      run.adjustment = TRUE
      if(FALSE){
        d3 <-density(Xa[,3]) 
        plot(d3)
        lines(d3$x,dnorm(d3$x,update$Y[3],1/sqrt(R[3,3])),col=2)
        lines(d3$x,dnorm(d3$x,priors$mufN[3],1/sqrt(priors$pfN[3,3])),col="blue")
        p.kf = R[3,3]+priors$pfN[3,3]
        mu.kf = priors$mufN[3]*priors$pfN[3,3]/p.kf+
          update$Y[3]*R[3,3]/p.kf
        lines(d3$x,dnorm(d3$x,mu.kf,1/sqrt(p.kf)),col=3,lty=2)

        jtest <- "model{
         Y ~ dmnorm(n[H],R[H,H])
        }"
        jtest <- "  ## SingleSite SingleDisturbance MultiPool
    model{
      ## prior
      n[3] ~ dnorm(mufN[3],pfN[3,3])  
      Y[H] ~ dnorm(n[H],R[H,H])
      dummy <- mufD[1] + pfD[1,1] + p
  }"
      jtest <- "  ## SingleSite SingleDisturbance MultiPool
    model{
      ## prior
      n[3] ~ dnorm(mufN[3],pfN[3,3])  
      Y[H] ~ dnorm(n[3],R[H,H])
  }"
        update2 = list(mufN = update$mufN,
                       pfN = update$pfN,
                       Y=update$Y,R=R,H=3)
        mtest <- rjags::jags.model(file=textConnection(jtest),
                                   data = update,
#                               data=list(n = c(NA,NA,mu.kf),
#                                         R=matrix(p.kf,3,3),H=3),
                               n.adapt=1000,n.chains=3,quiet=TRUE)
        mdat <- rjags::coda.samples(mtest,variable.names=c("n"),n.iter=10000) 
      summary(mdat)  
      }
    } else {
      Xa = Nf
      ca = 1
      run.adjustment = FALSE
    }

@@

## analyze sites - just one disturbance
```{r}
Rmult = c(1,0.1,0.01,0.001)
Qmult = c(1,10)
Pmult = c(1)
Ri = 4 #4
Qi = 1 #2
Pi = 1
#i = 181  # 177:263 = #fire
od = c(83:176,264:356)
for(i in 83:356){
  ofile = paste0("OR_SD_",
                 formatC(i, width = 3, format = "d", flag = "0"),
                 "_P",Pi,"Q",Qi,"R",Ri,".RData")
  print(ofile)
  NT = floor(ncol(rad)/365)
  Yi = matrix(NA,NT,3)  ## observed data
  Yi[1:ncol(Yw),3] = Yw[i,]  ## only Cw observed
  R = diag(rep(Rw,3))*Rmult[Ri]    ## observation error
  Q = diag(c(Ql,Qsoil,Qw)^2)*Qmult[Qi]  ## process error
  dtype = which(names(d.prior) == dextract[i,"t2004"])
  if(length(dtype)==0) break
  dist <- list(              ## disturbance parameters
    p.dist = d.prior[dtype]*Pmult[Pi],     ## disturbance probability
    mu0 = c(0.2,PBBAR[4,dtype]),   ## % retention [leaf,agb], first parameter made up, should include soil
    V0 =  diag(c(0.15,PBVAR[4,dtype])^2), #variance in retention
    alloc.soil = 0.05        ## made up  
  )
  test = SSSD.model(Y = Yi,R=R,Q=Q,
                  PAR=rad[i,],Xic=ICprior,THETA=theta[1:11],
                  dist = dist)
  save(test,file=ofile)
}
```

## visualization
```{r}
i = 181
pool.names=c("Cleaf","Csoil","Cstem")
for(i in 1:3){

@@

  plot(Yi[,i],
       ylim = range(rbind(CI.Na,Yi[,i]),na.rm=TRUE),
       type='n',
       ylab=pool.names[i],xlab="time")
  ecoforecastR::ciEnvelope(1:ncol(CI.Nf),CI.Nf[1,],CI.Nf[3,],col=ecoforecastR::col.alpha("blue",0.5))
  ecoforecastR::ciEnvelope(1:ncol(CI.Na),CI.Na[1,],CI.Na[5,],col=ecoforecastR::col.alpha('#b2df8a',0.7))
  lines(CI.Na[3,],col="#b2df8a",lwd=2)
  points(Yi[,i])
  lines(CI.D[3,])

@@

 TODO:
 * why is Analysis not contracting around the data?
    * reducing obs error by /10 didn't solve
 * why are the disturbance forecast pools crazy?
 * assess disturbance posterior

## Computer experiments
* Run under defaults
* Run under a grid of inc process error and decrease obs error and increase disturbance probability

@@

```{r}
## Detect set of experiments that have been run
runs <- dir(pattern="*.RData")
runs <- runs[grepl("^OR",runs)]
exp = sapply(strsplit(runs,"_",fixed=TRUE),function(x){
  sub(".RData","",x[4],fixed = TRUE)
})
exps <- unique(exp)

## load experiments that haven't yet been summarized
# elist <- list()   ## run once
t04 = 15            ## time of disturbance assessement (2004)
tlo = 28            ## time of last observation
for(i in seq_along(exps)){
  if(is.null(elist[[exps[i]]])){
    esel <- which(exp==exps[i])
    distrate = rep(NA,nrow(dextract))
    divergence = predist = distsize = matrix(NA,nrow(dextract),6) ## pre and post disturbance biomass pools (w/uncertainty)
    for(j in seq_along(esel)){
      ## load file
      load(runs[esel[j]])
      k = as.numeric(unlist(strsplit(runs[esel[j]],"_",fixed=TRUE))[3])

@@

    elist[[exps[i]]] = list(distrate=distrate,predist=predist,
                            distsize=distsize,divergence=divergence)

@@

save(elist,obs.pre,obs.dist,obs.tlo,obs.class,file="elist.RData")
```

@@

```{r}
pdf("OR_exps_diagnostics.pdf")
trim <- function(x,lo=0.0001,hi=0.9999){
  x[x<lo] <- lo
  x[x>hi] <- hi
  return(x)
}
betatrim <- function(y){
    n.obs <- sum(!is.na(y))
    (y * (n.obs - 1) + 0.5) / n.obs
}
stats <- list()
for(i in seq_along(exps)){

  ## divergence
  div = (elist[[i]]$divergence[,1] - obs.tlo)/
    (apply(elist[[i]]$divergence[,c(2,6)],1,diff)*0.5)
  div.reg    = lm(div~obs.tlo)
  div.median = median(div,na.rm = TRUE)
  div.cov    = sum(div < 1 & div > -1,na.rm=TRUE)/sum(!is.na(div))

@@

  ## Disturbance detection
  distrate = elist[[i]]$distrate
  hist(distrate,main=exps[i])

@@

  ## vs proportional disturbance size
  distprop = 1 - elist[[i]]$distsize[,1]/elist[[i]]$predist[,1] ## predicted
  distprop = 1 - obs.dist/obs.pre ## observed
  dist.beta.prop = betareg::betareg(distrate2 ~ distprop,
                                   na.action=na.omit,link="logit")
  plot(distprop,distrate,col=obs.class,main=exps[i])
  lines(pred.seq$distprop,predict(dist.beta.prop,newdata=pred.seq),col="green")

@@

}
dev.off()
save(stats,file="stats.RData")
```

factors affecting detection of disturbance
```{r}
Ps = as.numeric(substr(names(elist),2,2))
Qs = as.numeric(substr(names(elist),4,4))
Rs = as.numeric(substr(names(elist),6,6))

@@

div.slope = sapply(stats,function(x){coef(x$div.reg)[2]})
plot(Rmult[Rs],div.slope,col=Qs,log="x",xlab="Variance Multiplier",ylab="Divergance Slope")
## Disturbance @ default
summary(stats[[4]]$dist.combo)
## nothing matters

## Disturbance @ P1Q2R2 (intermediate)
summary(stats[[4]]$dist.combo)
## detection varies with divergence & disturbance type

## Disturbance magnitude: P1Q1R4
i = 3
summary(stats[[i]]$dist.combo)  ## absolute magnitude of divergence is best predictor
summary(stats[[i]]$dist.bin.abs)
beta = coef(stats[[i]]$dist.bin.abs)
## threshold mx+b = 0
thresh = -beta[1]/beta[2]

## Disturbance type: P1Q1R4 (reduced divergence)
stats[[i]]$dist.class
tapply(distabs,obs.class,summary) ## obs dist size by class
thresh.prop = tapply(distabs,obs.class,function(x){sum(x>thresh,na.rm = TRUE)/sum(!is.na(x))}) ## prop > detection threshold
thresh.prop
stats[[i]]$dist.class
dclass = stats[[i]]$dist.class
#dclass[1] = 0
dcol = as.numeric(names(stats[[i]]$dist.class)) + 1
plot(thresh.prop,dclass,col=dcol,pch=19)
abline(0,1,lty=2,col=2)
legend("topleft",legend = dcol-1,col=dcol,pch=19)

@@

## settings
site = 188
exp.name = "P1Q1R4" #"P1Q1R1"  ## manually run for each experiment **

@@

## model
load(paste0("OR_SD_",site,"_",exp.name,".RData"),envir = viz)
pool.names=c("Cleaf","Csoil","Cstem")
test = viz$test

@@

}
save(Yi,CI.Na,CI.D,file=paste0("OR_TS_",site,"_",exp.name,".RData"))
```

## summary info for filter divergence figures
```{r}
exp.name = "P1Q1R4" #"P1Q1R1"

@@

  ## divergence
  div = (elist[[i]]$divergence[,1] - obs.tlo)/
    (apply(elist[[i]]$divergence[,c(2,6)],1,diff)*0.5)
  div.reg    = lm(div~obs.tlo)
  div.median = median(div,na.rm = TRUE)
  div.cov    = sum(div < 1 & div > -1,na.rm=TRUE)/sum(!is.na(div))

>>>>

---
title: "OR Multi Disturbance"
author: "Michael Dietze"
date: "2024-05-27"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
source("disturbance.R")
source("multi.ens.adj.R")
library(BayesianTools)
```

## set up
```{r}
## Load extracted site timeseries
load("extract.h03v04.RData")

## Load disturbance rate, state&transition, disturbed biomass stats
load("STprior.RData")
dtime = 1985:2012
colnames(dextract) <- paste0("t",dtime)
dextract[is.na(dextract)] <- 0
d.prior["0"] = d.prior["0"] + d.prior[6]  ## change at runtime (removes <NA>)
d.prior = d.prior[-6]

## Load Calibration
load("fit.RData")
Qsoil = Qs
```

@@

## functions
```{r,echo=FALSE,results='hide',message=FALSE}
##' Single Site, Single Disturbance
##' @param Y observed data
##' @param R observation error
##' @param Q process error
##' @param Xic state initial condition matrix
##' @param THETA parameter matrix
##' @param dist  disturbance list: p.dist[1:NT], mu0[3], V0[3,3], alloc soil 
SSSD.model <- function(Y,R,Q,PAR,Xic,THETA,dist){

@@

    ## Disturbance
    d.forecast <- rep(0,ne); d.forecast[1:ceiling(ne*0.1)] <- 1 #rbinom(ne,1,dist$p.dist)             ## prior on disturbance status
    sel <- !as.logical(d.forecast)                ## select undisturbed ensemble members
    if(sum(!sel)>2){
      sel.dist = which(!sel)
      for(j in sel.dist){                      ## run disturbance model
        Nf[j,]  = disturbance.p(Nf[j,],mu0=dist$mu0,V0 = dist$V0,alloc.soil = dist$alloc.soil)
      }
      mufD  = colMeans(Nf[!sel,])
      pfD = solve(var(Nf[!sel,]))#solve(diag(c(0.15,0.5,0.25)^2))
    } 
    if(sum(!sel) <= 2 || "try-error" %in% class(pfD)) {   ## backup values if the ensemble stats fail
      mufD = c(dist$mu0[1], mean(Nf[,2]),  dist$mu0[2])            
      pfD  = solve(diag(c(0.15,0.5,0.25)^2))
    }
    Nf <- tobit(Nf)

    ## priors
    priors <- list(p    = dist$p.dist[t],         ## disturbance rate
                   mufN = colMeans(Nf[sel,]),    ## Not disturbed forecast means
                   mufD = mufD,   ## Disturbed forecast means
                   pfN  = solve(var(Nf[sel,])),  ## Not disturbed forecast precision
                   pfD  = pfD  ## Disturbed forecast precision
    )
    Forecast[[t]] <- c(list(d.forecast=d.forecast,Nf=Nf),priors)

@@

      ## update parameters & save Analysis posterior
      dat = as.matrix(jdat)
      nsel <- grep(pattern="^n",x=colnames(dat))
      Xa = dat[,nsel]
      ca = dat[,"D"]
      run.adjustment = TRUE
    } else {
      Xa = Nf
      ca = 1
      run.adjustment = FALSE
    }

@@

## analyze sites - just one disturbance
```{r}
Rmult = c(1,0.1,0.01,0.001)
Qmult = c(1,10)
Pmult = c(1)
Ri = 2
Qi = 1
Pi = 1
users_accuracy = c(1-d.prior[1],.77,.92,.77,.78,.78) #Kennedy et al 2015 table 6
## 0 = none, 10 = other, 20 = cut, 30 = road, 40 = fire, 50 = biotic
od = c(83:176,264:356)
for(i in 1:356){
  ofile = paste0("OR_DP_",  ## DP = Disturbance Prior
                 formatC(i, width = 3, format = "d", flag = "0"),
                 "_P",Pi,"Q",Qi,"R",Ri,".RData")
  print(ofile)
  if(file.exists(ofile)) next
  NT = floor(ncol(rad)/365)
  Yi = matrix(NA,NT,3)  ## observed data
  Yi[1:ncol(Yw),3] = Yw[i,]  ## only Cw observed
  R = diag(rep(Rw,3))*Rmult[Ri]    ## observation error
  Q = diag(c(Ql,Qsoil,Qw)^2)*Qmult[Qi]  ## process error
  dtype = which(names(d.prior) == dextract[i,"t2004"])
  if(dtype == 1) dtype = sample(2:5,1) ## if site is undisturbed, set alternative at random
  if(length(dtype)==0) break
  p.dist = rep(d.prior[dtype]*Pmult[Pi],NT) ## set background rate
  has.dist = which(dextract[i,] > 0)
  p.dist[has.dist] = users_accuracy[dextract[i,has.dist]/10 + 1] 
  dist <- list(              ## disturbance parameters
    #p.dist = d.prior[dtype]*Pmult[Pi],     ## disturbance probability
    p.dist = p.dist,
    mu0 = c(0.2,PBBAR[4,dtype]),   ## % retention [leaf,agb], first parameter made up, should include soil
    V0 =  diag(c(0.15,PBVAR[4,dtype])^2), #variance in retention
    alloc.soil = 0.05        ## made up  
  )
  test = SSSD.model(Y = Yi,R=R,Q=Q,
                  PAR=rad[i,],Xic=ICprior,THETA=theta[1:11],
                  dist = dist)
  save(test,dist,file=ofile)
  test$Analysis[[20]]$mu
}
```

## visualization
```{r}
i = 188#181
ofile = paste0("OR_DP_",  ## DP = Disturbance Prior
                 formatC(i, width = 3, format = "d", flag = "0"),
                 "_P",Pi,"Q",Qi,"R",Ri,".RData")
load(ofile)
NT = floor(ncol(rad)/365)
Yi = matrix(NA,NT,3)  ## observed data
Yi[1:ncol(Yw),3] = Yw[i,]  ## only Cw observed
pool.names=c("Cleaf","Csoil","Cstem")
for(i in 1:3){

@@

  plot(Yi[,i],
       ylim = range(c(range(rbind(CI.Na,Yi[,i]),na.rm=TRUE),0)),
       type='n',
       ylab=pool.names[i],xlab="time")
  ecoforecastR::ciEnvelope(1:ncol(CI.Nf),CI.Nf[1,],CI.Nf[3,],col=ecoforecastR::col.alpha("blue",0.5))
  ecoforecastR::ciEnvelope(1:ncol(CI.Na),CI.Na[1,],CI.Na[5,],col=ecoforecastR::col.alpha('#b2df8a',0.7))
  lines(CI.Na[3,],col="#b2df8a",lwd=2)
  points(Yi[,i])
  lines(CI.D[3,])

@@

 TODO:
 * why is filter divergence such a problem even prior to disturbance and with prescribed disturbance probabilities?

@@

```{r}
## Detect set of experiments that have been run
runs <- dir(pattern="*.RData")
runs <- runs[grepl("^OR_DP",runs)]
exp = sapply(strsplit(runs,"_",fixed=TRUE),function(x){
  sub(".RData","",x[4],fixed = TRUE)
})
exps <- unique(exp)

## load experiments that haven't yet been summarized
# eDPlist <- list()   ## run once
t04 = 15            ## time of disturbance assessement (2004)
tlo = 28            ## time of last observation
for(i in seq_along(exps)){
  if(is.null(eDPlist[[exps[i]]])){
    esel <- which(exp==exps[i])
    distrate = rep(NA,nrow(dextract))
    divergence = predist = distsize = matrix(NA,nrow(dextract),6) ## pre and post disturbance biomass pools (w/uncertainty)
    for(j in seq_along(esel)){
      ## load file
      load(runs[esel[j]])
      k = as.numeric(unlist(strsplit(runs[esel[j]],"_",fixed=TRUE))[3])

@@

    eDPlist[[exps[i]]] = list(distrate=distrate,predist=predist,
                            distsize=distsize,divergence=divergence)

@@

save(eDPlist,obs.pre,obs.dist,obs.tlo,obs.class,file="eDPlist.RData")
```

@@

```{r}
pdf("OR_DP_exps_diagnostics.pdf")
trim <- function(x,lo=0.0001,hi=0.9999){
  x[x<lo] <- lo
  x[x>hi] <- hi
  return(x)
}
betatrim <- function(y){
    n.obs <- sum(!is.na(y))
    (y * (n.obs - 1) + 0.5) / n.obs
}
stats <- list()
for(i in seq_along(exps)){

  ## divergence
  div = (eDPlist[[i]]$divergence[,1] - obs.tlo)/
    (apply(eDPlist[[i]]$divergence[,c(2,6)],1,diff)*0.5)
  div.reg    = lm(div~obs.tlo)
  div.median = median(div,na.rm = TRUE)
  div.cov    = sum(div < 1 & div > -1,na.rm=TRUE)/sum(!is.na(div))

@@

  ## Disturbance detection
  distrate = eDPlist[[i]]$distrate
  hist(distrate,main=exps[i])

@@

  ## vs proportional disturbance size
  distprop = 1 - eDPlist[[i]]$distsize[,1]/eDPlist[[i]]$predist[,1] ## predicted
  distprop = 1 - obs.dist/obs.pre ## observed
  dist.beta.prop = betareg::betareg(distrate2 ~ distprop,
                                   na.action=na.omit,link="logit")
  plot(distprop,distrate,col=obs.class,main=exps[i])
  lines(pred.seq$distprop,predict(dist.beta.prop,newdata=pred.seq),col="green")

@@

}
dev.off()
save(stats,file="stats.DP.RData")
```
# TODO
* add figures looking at how AGB data updates disturbance probability relative to the priors

factors affecting detection of disturbance
```{r}
Ps = as.numeric(substr(names(eDPlist),2,2))
Qs = as.numeric(substr(names(eDPlist),4,4))
Rs = as.numeric(substr(names(eDPlist),6,6))

@@

div.slope = sapply(stats,function(x){coef(x$div.reg)[2]})
plot(Rmult[Rs],div.slope,col=Qs,log="x",xlab="Variance Multiplier",ylab="Divergance Slope")
## Disturbance @ default
#summary(stats[[4]]$dist.combo)
## nothing matters

## Disturbance @ P1Q2R2 (intermediate)
#summary(stats[[4]]$dist.combo)
## detection varies with divergence & disturbance type

## Disturbance magnitude: P1Q1R2
i = 2
summary(stats[[i]]$dist.combo)  ## absolute magnitude of divergence is best predictor
summary(stats[[i]]$dist.bin.abs)
beta = coef(stats[[i]]$dist.bin.abs)
## threshold mx+b = 0
thresh = -beta[1]/beta[2]

## Disturbance type: P1Q1R2 (reduced divergence)
stats[[i]]$dist.class
tapply(distabs,obs.class,summary) ## obs dist size by class
thresh.prop = tapply(distabs,obs.class,function(x){sum(x>thresh,na.rm = TRUE)/sum(!is.na(x))}) ## prop > detection threshold
thresh.prop
stats[[i]]$dist.class
dclass = stats[[i]]$dist.class
#dclass[1] = 0
dcol = as.numeric(names(stats[[i]]$dist.class)) + 1
plot(thresh.prop,dclass,col=dcol,pch=19)
abline(0,1,lty=2,col=2)
legend("topleft",legend = dcol-1,col=dcol,pch=19)

@@

## settings
site = 188
exp.name = "P1Q1R2" #"P1Q1R1"  ## manually run for each experiment **

@@

## model
viz = rlang::env()
load(paste0("OR_DP_",site,"_",exp.name,".RData"),envir = viz)
pool.names=c("Cleaf","Csoil","Cstem")
test = viz$test

@@

}
save(Yi,CI.Na,CI.D,file=paste0("OR_DPTS_",site,"_",exp.name,".RData"))
```

## summary info for filter divergence figures
```{r}
exp.name = "P1Q1R2" #"P1Q1R1"

@@

  ## divergence
  div = (eDPlist[[i]]$divergence[,1] - obs.tlo)/
    (apply(eDPlist[[i]]$divergence[,c(2,6)],1,diff)*0.5)
  div.reg    = lm(div~obs.tlo)
  div.median = median(div,na.rm = TRUE)
  div.cov    = sum(div < 1 & div > -1,na.rm=TRUE)/sum(!is.na(div))
